\documentclass[../../main.tex]{subfiles} % Use the main document's preamble

\begin{document}
\chapter{MOTs}

https://journals.aps.org/pra/pdf/10.1103/PhysRevA.70.063413 narrow line cooling and momentum space crystals

https://journals.aps.org/prl/pdf/10.1103/PhysRevLett.93.073003 previous paper narrow line cooling finite photon recoil dynamics
For the 689 line in strontium, the ratio of the recoil frequency $\omega_r=\frac{\hbar k^2}{2m}=2\pi\times 4.8$ kHz to the natural linewidth $\Gamma=2\pi\times 7.4$ kHz the transition is on the order of unity. this is what defines a narrow line MOT.
\section{Finite photon recoil dynamics}
The 1s0-3p1 transition has its single photon recoil is of comparable sixe to the natural linewidth. When an atom of mass $m$ emits or absorbs a photon $\gamma$ of frequency $\omega$, then momentum of the photon $p_\gamma=h/\lambda=\hbar k$ is transfered to the atom giving it some velocity $v_{rec}=\hbar k/m$



\section{Regimes of red-detuned MOT}
Three regimes of behavior depend on the relative magnitudes of $|\Delta|$, $\Gamma$, and $\Gamma_E=\Gamma\sqrt{1+s}$ where $\Gamma$ is the natural linewidth $s=I/I_{sat}$ is the saturation intensity which may broaden the line. The three regimes are (I) $\Gamma\ll\Gamma_E\ll |\Delta|$, (II), $|\Delta|<\Gamma_E$ and $\Gamma\ll\Gamma_E$ so $\Gamma<|\Delta|<\Gamma_e$ (but perhaps $\Gamma\centernot\ll |\Delta|$?), and (III), when $s\simeq 1$. and so $\Gamma\centernot\ll \Gamma_E$. 

\subsection{Red-detuned MOT Regime (I)}
In the first regime, $\Gamma\ll\Gamma_E\ll |\Delta|$. For the 689 line in strontium, the ratio of the recoil frequency $\omega_r=\frac{\hbar k^2}{2m}=2\pi\times 4.8$ kHz to the natural linewidth $\Gamma=2\pi\times 7.4$ kHz the transition is on the order of unity. In this regime, 

\subsection{regime 2 }
$|\Delta| \lesssim \Gamma_E$ and $\Gamma\ll\Gamma_E$ 

this is more typical and thermodynamics remicent of ordinary doppler cooling. $\delta$ and $s$ dependent teperature minima

\subsection{regime 3}
s approces unit, photon-tecoil-driven impulsive force dominates and temperature below photon-recoil limit $T_r=2\hbar\omega_r/k_b$

$$\Gamma\simeq\Gamma_E$$

\subsection{other regimes?}
sing $\Gamma\sim \omega_r$ you get momentum packets. explained by analytic solution to the 1D semiclassical radiative force equation. for $\delta>0$, and $\Gamma\sim \omega_r$  allow direct visualization of positive feedback acceleration that effeciently bunches the atoms into discrete, well-defined momentum packets. the gravity ratio $R$ directly influences the blue detuned thermodynamics, enabling cooling around a velocyt where radiation pressure and gravity balance. aparently the underlying physics here is the same as regime I but manifest bvery differently.

\subsection{method}
uses frequency modulate redMOT, start at 3G/cm and go to 10 G/cm. Then transition to singlefrequency 


\end{document}